---
title: "KadenFlorsheimDavidLüdeke_Final"
author: "Kaden Florsheim"
date: "11/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(mapview)
library(leaflet)
library(censusapi)
library(gtools)
Sys.setenv(CENSUS_KEY="9fbd5ddd430b595b8f3715733cae2b75c18be92e")
```

```{r}
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO", #Unique ID for each household
    "SPORDER", #Person number
    "PWGTP", #Total number of people
    "WGTP", #Housing Weight
    "HINCP", #Household Income
    "CPLT", #Couple Type
    "FINCP", #Family Income
    "PINCP", #Total Persons Income
    "RAC1P" #Recorded Detailed Race Code
  )
)
```

```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

counties <-
  counties("CA", cb = T, progress_bar = F)

ca_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

ca_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% ca_pumas$PUMACE10)
```

```{r}
cleaned <- ca_pums %>%
  mutate(
    SPORDER = as.numeric(SPORDER),
    CPLT = as.numeric(CPLT)
  ) %>% 
  filter(
    (SPORDER %in% 1:2), #Only using head of household data so that household income is not double counted thus skewing results.
    (CPLT %in% 1:4)) %>% #filtering out N/A
  select(-public_use_microdata_area)
```

```{r}
compare <-
  cleaned %>% 
  pivot_wider(
    names_from = "SPORDER",
    values_from = c("RAC1P"))
```


## EQUITY ANALYSIS BETWEEN SAME-SEX COUPLES AND OPPOSITE-SEX COUPLES DISAGGREGATED BY RACE 
```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    RAC1P = as.numeric(RAC1P),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), #Same-sex married couple|same-sex unmarried couple
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 20000),
       "$20,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 20001 & HINCP <= 50000),
       "$20,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50001 & HINCP <= 80000),
       "$50,001 to $80,000",
      income
    ),
    income = ifelse(
      (HINCP > 80001 & HINCP <= 100000),
       "$80,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100001 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150001 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200001 & HINCP <= 250000),
       "$200,001 to $250,000",
      income
    ),
    income = ifelse(
      (HINCP > 250001),
       "$250,001 or more",
      income
    ))
      
      
    #   if(HINCP < 10000){
    #   "Less than $10,000"
    # } else if (HINCP > 10001 & HINCP > 14999){
    #   "10,000 to 14,999"
    # } else if (HINCP > 15000 & HINCP > 19999){
    #   "15,000 to 19,999"
    # } else if (HINCP > 20000 & HINCP > 24999){
    #   "20,000 to 24,999"
    # } else if (HINCP > 25000 & HINCP > 49000){
    #   "25,000 to 49,000"
    # } else {
    #   "More than 50,000"}
    # )

```  

```{r}
order <- c( "$20,000 or less", "$20,001 to $50,000","$50,001 to $80,000","$80,001 to $100,000","$100,001 to $150,000","$150,001 to $200,000","$200,001 to $250,000","$250,001 or more")
```
 
```{r}
ca_race_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_race_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), 
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 20000),
       "$20,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 20001 & HINCP <= 50000),
       "$20,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50001 & HINCP <= 80000),
       "$50,001 to $80,000",
      income
    ),
    income = ifelse(
      (HINCP > 80001 & HINCP <= 100000),
       "$80,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100001 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150001 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200001 & HINCP <= 250000),
       "$200,001 to $250,000",
      income
    ),
    income = ifelse(
      (HINCP > 250001),
       "$250,001 or more",
      income
    ),
    RAC1P = as.numeric(RAC1P)
  ) %>% 
  filter(
    (RAC1P %in% 1)) #white alone
```

## White Alone
```{r}
ca_white_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_white_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), 
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 20000),
       "$20,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 20001 & HINCP <= 50000),
       "$20,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50001 & HINCP <= 80000),
       "$50,001 to $80,000",
      income
    ),
    income = ifelse(
      (HINCP > 80001 & HINCP <= 100000),
       "$80,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100001 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150001 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200001 & HINCP <= 250000),
       "$200,001 to $250,000",
      income
    ),
    income = ifelse(
      (HINCP > 250001),
       "$250,001 or more",
      income
    ),
    RAC1P = as.numeric(RAC1P)
  ) %>% 
  filter(
    (RAC1P %in% 2)) #Black or African American alone
```

## Black or African American alone
```{r}
ca_black_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_black_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), 
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 10000),
       "$10,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 10000 & HINCP <= 20000),
       "$10,001 to $20000",
      income
    ),
    income = ifelse(
      (HINCP > 20000 & HINCP <= 30000),
       "$20,001 to $30,000",
      income
    ),
    income = ifelse(
      (HINCP > 30000 & HINCP <= 40000),
       "$30,001 to $40,000",
      income
    ),
    income = ifelse(
      (HINCP > 40000 & HINCP <= 50000),
       "$40,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50000 & HINCP <= 75000),
       "$50,001 to $75,000",
      income
    ),
    income = ifelse(
      (HINCP > 75000 & HINCP <= 100000),
       "$75,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100000 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150000 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200000),
       "$200,001 or more",
      income
    ),
    RAC1P = as.numeric(RAC1P)
  ) %>% 
  filter(
    (RAC1P %in% 6)) #Asian American alone
```


```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), 
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 10000),
       "$10,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 10000 & HINCP <= 20000),
       "$10,001 to $20000",
      income
    ),
    income = ifelse(
      (HINCP > 20000 & HINCP <= 30000),
       "$20,001 to $30,000",
      income
    ),
    income = ifelse(
      (HINCP > 30000 & HINCP <= 40000),
       "$30,001 to $40,000",
      income
    ),
    income = ifelse(
      (HINCP > 40000 & HINCP <= 50000),
       "$40,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50000 & HINCP <= 75000),
       "$50,001 to $75,000",
      income
    ),
    income = ifelse(
      (HINCP > 75000 & HINCP <= 100000),
       "$75,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100000 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150000 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200000),
       "$200,001 or more",
      income
    ),
    RAC1P = as.numeric(RAC1P)
  ) %>% 
  filter(
    (RAC1P %in% 3)) #American Indian alone
```


## American Indian alone
```{r}
ca_AI_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_AI_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```





## Asian American Alone
```{r}
ca_asian_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_asian_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

```{r}
ca_pums_IRC <-
  cleaned %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    Same_sex_household = ifelse(
        (CPLT == 2)|(CPLT == 4), 
      "Same sex couple",
      "Opposite sex couple"
    ),
    income =  ifelse(
      (HINCP <= 10000),
       "$10,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 10000 & HINCP <= 20000),
       "$10,001 to $20000",
      income
    ),
    income = ifelse(
      (HINCP > 20000 & HINCP <= 30000),
       "$20,001 to $30,000",
      income
    ),
    income = ifelse(
      (HINCP > 30000 & HINCP <= 40000),
       "$30,001 to $40,000",
      income
    ),
    income = ifelse(
      (HINCP > 40000 & HINCP <= 50000),
       "$40,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50000 & HINCP <= 75000),
       "$50,001 to $75,000",
      income
    ),
    income = ifelse(
      (HINCP > 75000 & HINCP <= 100000),
       "$75,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100000 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150000 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200000),
       "$200,001 or more",
      income
    ),
    RAC1P = as.numeric(RAC1P)
  ) %>% 
  filter(
    (RAC1P %in% 9)) #Two or More Races
```

## Two or More Races
```{r}
ca_two_total <-
  ca_pums_IRC %>% 
  group_by(Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, Same_sex_household) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_two_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = Same_sex_household %>% factor(levels = rev(unique(ca_pums_IRC$Same_sex_household)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California household income by couple type",
    fill = "Couple Type"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
``` 



## SAME-SEX AND RACE EQUITY ANALYSIS

```{r}
ca_pums_IRC <-
  cleaned %>%
  filter(
    (CPLT %in% c(2, 4)), #Filtering out opposite-sex couples
    (RAC1P %in% c(1,2,3,6,9))) %>%  #Filtering out RAC1P variables that don't have enough data (Alaska Native alone,American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races, Native Hawaiian and Other Pacific Islander alone, Some Other Race alone)
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    income =  ifelse(
      (HINCP <= 10000),
       "$10,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 10000 & HINCP <= 20000),
       "$10,001 to $20000",
      income
    ),
    income = ifelse(
      (HINCP > 20000 & HINCP <= 30000),
       "$20,001 to $30,000",
      income
    ),
    income = ifelse(
      (HINCP > 30000 & HINCP <= 40000),
       "$30,001 to $40,000",
      income
    ),
    income = ifelse(
      (HINCP > 40000 & HINCP <= 50000),
       "$40,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50000 & HINCP <= 75000),
       "$50,001 to $75,000",
      income
    ),
    income = ifelse(
      (HINCP > 75000 & HINCP <= 100000),
       "$75,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100000 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150000 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200000),
       "$200,001 or more",
      income
    ),
    race = ifelse(
      (RAC1P == 1),
      "White Only",
      RAC1P
    ),
    race = ifelse(
      (RAC1P == 2),
      "Black or African American alone",
      race
    ),
    race = ifelse(
      (RAC1P == 3),
      "American Indian alone",
      race
    ),
    race = ifelse(
      (RAC1P == 6),
      "Asian alone",
      race
    ),
    race = ifelse(
      (RAC1P == 9),
      "Two or More Races",
      race
    )) 
```

```{r}
ca_race_total <-
  ca_pums_IRC %>% 
  group_by(race) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_race_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(ca_pums_IRC$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California same-sex household income by Race",
    fill = "Race"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

```{r}
ca_pums_IRC <-
  cleaned %>%
  filter(
    (CPLT %in% c(1, 3)), #Filtering out same-sex couples
    (RAC1P %in% c(1,2,3,6,9))) %>%  #Filtering out RAC1P variables that don't have enough data (Alaska Native alone,American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races, Native Hawaiian and Other Pacific Islander alone, Some Other Race alone)
  mutate(
    PWGTP = as.numeric(PWGTP),
    HINCP = as.numeric(HINCP),
    income =  ifelse(
      (HINCP <= 10000),
       "$10,000 or less",
      HINCP
    ),
    income = ifelse(
      (HINCP > 10000 & HINCP <= 20000),
       "$10,001 to $20000",
      income
    ),
    income = ifelse(
      (HINCP > 20000 & HINCP <= 30000),
       "$20,001 to $30,000",
      income
    ),
    income = ifelse(
      (HINCP > 30000 & HINCP <= 40000),
       "$30,001 to $40,000",
      income
    ),
    income = ifelse(
      (HINCP > 40000 & HINCP <= 50000),
       "$40,001 to $50,000",
      income
    ),
    income = ifelse(
      (HINCP > 50000 & HINCP <= 75000),
       "$50,001 to $75,000",
      income
    ),
    income = ifelse(
      (HINCP > 75000 & HINCP <= 100000),
       "$75,001 to $100,000",
      income
    ),
    income = ifelse(
      (HINCP > 100000 & HINCP <= 150000),
       "$100,001 to $150,000",
      income
    ),
    income = ifelse(
      (HINCP > 150000 & HINCP <= 200000),
       "$150,001 to $200,000",
      income
    ),
    income = ifelse(
      (HINCP > 200000),
       "$200,001 or more",
      income
    ),
    race = ifelse(
      (RAC1P == 1),
      "White Only",
      RAC1P
    ),
    race = ifelse(
      (RAC1P == 2),
      "Black or African American alone",
      race
    ),
    race = ifelse(
      (RAC1P == 3),
      "American Indian alone",
      race
    ),
    race = ifelse(
      (RAC1P == 6),
      "Asian alone",
      race
    ),
    race = ifelse(
      (RAC1P == 9),
      "Two or More Races",
      race
    )) 
```

```{r}
ca_race_total <-
  ca_pums_IRC %>% 
  group_by(race) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  mutate(income = "Total")

ca_pums_IRC %>% 
  group_by(income, race) %>% 
  summarize(estimate = sum(PWGTP)) %>% 
  rbind(ca_race_total) %>%  
  ggplot() +
  geom_bar(
    aes(
      x = income %>% factor(levels = rev(c("Total", mixedsort(order)))),
      y = estimate,
      fill = race %>% factor(levels = rev(unique(ca_pums_IRC$race)))
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Household income",
    y = "Number of households",
    title = "California opposite-sex household income by race",
    fill = "Race"
  ) +
  coord_flip() +
  theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )
```

















# 
# ```{r}
# census_race_categories <-
#   c(
#       "White alone",
#       "Black or African American alone",
#       "American Indian alone",
#       "Alaska Native alone",
#       "American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races",
#        "Asian alone",
#        "Native Hawaiian and Other Pacific Islander alone",
#        "Some Other Race alone",
#        "Two or More Races")
# ```
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# ```{r}
# ca_income_race <- 
#   1:9 %>% 
#   map_dfr(function(x){
#     race <- cleaned %>%
#   mutate(
#     RAC1P = as.numeric(RAC1P)
#   ) %>% 
#   filter(
#     (RAC1P %in% census_race_categories[x])) %>% 
#   mutate(
#     PWGTP = as.numeric(PWGTP),
#     HINCP = as.numeric(HINCP),
#     Same_sex_household = ifelse(
#         (CPLT == 2)|(CPLT == 4), 
#       "Same sex couple",
#       "Opposite sex couple"
#     ),
#     income =  ifelse(
#       (HINCP <= 10000),
#        "$10,000 or less",
#       HINCP
#     ),
#     income = ifelse(
#       (HINCP > 10000 & HINCP <= 20000),
#        "$10,001 to $20000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 20000 & HINCP <= 30000),
#        "$20,001 to $30,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 30000 & HINCP <= 40000),
#        "$30,001 to $40,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 40000 & HINCP <= 50000),
#        "$40,001 to $50,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 50000 & HINCP <= 75000),
#        "$50,001 to $75,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 75000 & HINCP <= 100000),
#        "$75,001 to $100,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 100000 & HINCP <= 150000),
#        "$100,001 to $150,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 150000 & HINCP <= 200000),
#        "$150,001 to $200,000",
#       income
#     ),
#     income = ifelse(
#       (HINCP > 200000),
#        "$200,001 or more",
#       income
#     )) %>% 
#   group_by(income, Same_sex_household) %>% 
#   summarize(estimate = sum(PWGTP)) %>% 
#   ggplot() +
#   geom_bar(
#     aes(
#       x = income %>% factor(mixedsort(order)),
#       y = estimate,
#       fill = Same_sex_household
#     ),
#     stat = "identity",
#     position = "stack"
#   ) +
#   labs(
#     x = "Household income",
#     y = "Number of households",
#     title = "Bay Area household income by couple type",
#     fill = "Couple Type"
#   ) +
#   coord_flip() +
#   theme(
#     legend.position = "bottom",
#     legend.direction = "vertical"
#   )  +
#   guides(
#     fill = guide_legend(
#       reverse = T
#     )
#   )
#   })
# 
# ```          
#           
#           
#           
#           
#     getCensus(
#       name = "acs/acs1/pums",
#       vintage = 2019,
#       region = "public use microdata area:*", 
#       regionin = "state:06",
#       vars = c(
#         "SERIALNO", #Unique ID for each household
#         "SPORDER", #Person number
#         "PWGTP", #Total number of people
#         "WGTP", #Housing Weight
#         "HINCP", #Household Income
#         "CPLT", #Couple Type
#         "FINCP", #Family Income
#         "PINCP", #Total Persons Income
#         "RAC1P" #Recorded Detailed Race Code
#       )
#     ) %>%
#       select(!c(GEO_ID,state,NAME) & !ends_with(c("EA","MA","M"))) %>%
#       pivot_longer(
#         ends_with("E"),
#         names_to = "name",
#         values_to = "estimate"
#       ) %>%
#       left_join(
#         ca_pums %>% 
#           select(name, label)
#       ) %>% 
#       select(-name) %>% 
#       separate(
#         label,
#         into = c(NA,NA,"income"),
#         sep = "!!"
#       ) %>% 
#       filter(!is.na(income)) %>% 
#       mutate(race = census_race_categories[x])
#   })
# ```
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
#       (HINCP < 10000),
#       "Less than 10,000",
#       (HINCP > 10000),
#       "More than 10,000",
#       0
#     ))
#     
#    
#     
#     
#     # IR = ifelse(
#     #   SERIALNO == SERIALNO|RAC1P!=RAC1P,
#     #   1,
#     #   0
#     # )
#     # )
#     # 
#     # 
# 
# # IRC_compare <-   #Attempting to pull out Interracial couples into new data set so we can compare races
# #   ca_pums_IRC %>% 
# #   mutate(
# #     PWGTP = as.numeric(PWGTP),
# #     SERIALNO = as.numeric(SERIALNO)) %>% 
# #   if(SERIALNO == SERIALNO) {
# #     mutate(
# #     PWGTP = as.numeric(PWGTP),
# #     Same_sex_household = ifelse(
# #         (CPLT == 2)|(CPLT == 4), 
# # 
# #       1,
# #       0
# #     )) 
# #   }
# ##Can't figure out how to separate interracial couples--Trying to do if statement
# #if {same serialno but different races put them in this data set or create new column}
# ```
#  
# ```{r}
# ca_pums_IRC %>% 
#   group_by(HINCP, Same_sex_household) %>% 
#   summarize(estimate = sum(PWGTP)) %>% 
#   ggplot() +
#   geom_bar(
#     aes(
#       x = HINCP,
#       y = estimate,
#       fill = Same_sex_household
#     ),
#     stat = "identity",
#     position = "stack"
#   ) +
#   labs(
#     x = "Household income",
#     y = "Number of households",
#     title = "Bay Area household income by couple type",
#     fill = "Couple Type"
#   )
# ``` 
#  
#  
#  
#  
#     
# ```{r}
# census_race_categories <- data.frame(
#   code = c(1,2,3,4,5,6,7,8,9),
#   category =
#     c(
#       "White alone",
#       "Black or African American alone",
#       "American Indian alone",
#       "Alaska Native alone",
#       "American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races",
#        "Asian alone",
#        "Native Hawaiian and Other Pacific Islander alone",
#        "Some Other Race alone",
#        "Two or More Races"
#     ))
# ```
#   
# ```{r}
# census_race_categories <-
#   c(
#       "White alone",
#       "Black or African American alone",
#       "American Indian alone",
#       "Alaska Native alone",
#       "American Indian and Alaska Native tribes specified; or American Indian or Alaska Native, not specified and no other races",
#        "Asian alone",
#        "Native Hawaiian and Other Pacific Islander alone",
#        "Some Other Race alone",
#        "Two or More Races")
# ```
# 
# ```{r}
# ca_income_race <- 
#   1:9 %>% 
#   map_dfr(function(x){
#     getCensus(
#       name = "acs/acs1/pums",
#       vintage = 2019,
#       region = "public use microdata area:*", 
#       regionin = "state:06",
#       vars = c(
#         "SERIALNO", #Unique ID for each household
#         "SPORDER", #Person number
#         "PWGTP", #Total number of people
#         "WGTP", #Housing Weight
#         "HINCP", #Household Income
#         "CPLT", #Couple Type
#         "FINCP", #Family Income
#         "PINCP", #Total Persons Income
#         "RAC1P" #Recorded Detailed Race Code
#       )
#     ) %>%
#       select(!c(GEO_ID,state,NAME) & !ends_with(c("EA","MA","M"))) %>%
#       pivot_longer(
#         ends_with("E"),
#         names_to = "name",
#         values_to = "estimate"
#       ) %>%
#       left_join(
#         ca_pums %>% 
#           select(name, label)
#       ) %>% 
#       select(-name) %>% 
#       separate(
#         label,
#         into = c(NA,NA,"income"),
#         sep = "!!"
#       ) %>% 
#       filter(!is.na(income)) %>% 
#       mutate(race = census_race_categories[x])
#   })
# ```
#  
# ```{r}
# ca_cup <-
#   getCensus(
#       name = "acs/acs1/pums",
#       vintage = 2019,
#       region = "public use microdata area:*", 
#       regionin = "state:06",
#       vars = c(
#         "SERIALNO", #Unique ID for each household
#         "SPORDER", #Person number
#         "PWGTP", #Total number of people
#         "WGTP", #Housing Weight
#         "HINCP", #Household Income
#         "CPLT", #Couple Type
#         "FINCP", #Family Income
#         "PINCP", #Total Persons Income
#         "RAC1P") #Recorded Detailed Race Code
#   ) %>%
#   left_join(
#     ca_pums_IRC %>% 
#       select(SERIALNO, public_use_microdata_area)
#   ) 
# 
# ```
# %>% 
#   select(-name) %>% 
#   separate(
#     label,
#     into = c(NA,NA,"income"),
#     sep = "!!"
#   ) %>% 
#   filter(!is.na(income))
# ```  
#   
#   ,
#     White_Low_Income = ifelse(
#       (RAC1P == 1) & (HINCP > 100000),
#       1,
#       0
#     ),
#     Not_White_Low_Income = ifelse(
#       (RAC1P != 1) & (HINCP > 100000),
#       1,
#       0)) 
# ```
# 
# ```{r}
# bay_pums_couple <-
#   cleaned %>% 
#   Same_sex_household = ifelse(
#         (CPLT == 2)|(CPLT == 4), 
#         WGTP,
#         0)
# # %>% 
#   # group_by(PUMA) %>% 
#   # summarize(
#   #   PERSON1 =
#   #     sum(Person1, na.rm =T)
#   # ) 
# ```
# ```
# 
# ```{r}
# bay_pums_couple <-
#   cleaned %>%
#   mutate(
#     WGTP = as.numeric(WGTP),
#     SPORDER = as.numeric(SPORDER),
#     partner1 = ifelse(
#       (SPORDER == 1),
#       WGTP,
#       0
#     ),
#     partner2 = ifelse(
#       (SPORDER == 2),
#       WGTP,
#       0
#     ) %>% 
#   group_by(PUMA) %>%
#   summarize(
#     partner1 =
#       sum(partner1, na.rm =T),
#     partner2 =
#       sum(partner2, na.rm =T))
#   )
# ```
# 
# 
# 
#     
#     ,
#     Same_sex_household = ifelse(
#       (HINCP < 100000) &
#         (CPLT == 2)|(CPLT == 4), 
#         
#       WGTP,
#       0
#     ),
#     Opposite_sex_household = ifelse(
#       (HINCP < 100000) &
#         (CPLT == 1)|(CPLT == 3),
#       WGTP,
#       0
#     ),
#     White = ifelse(
#       (SERIALNO == 1) &
#       (RAC1P == 1),
#       WGTP,
#       0
#     )
#     
#     
#     ) %>% 
#   group_by(PUMA) %>%
#   summarize(
#     sum_same_sex_household =
#       sum(Same_sex_household, na.rm =T),
#     sum_opposite_sex_household =
#       sum(Opposite_sex_household, na.rm =T),
#     partner1 =
#       sum(partner1, na.rm =T),
#     partner2 =
#       sum(partner2, na.rm =T)%>% 
#   left_join(
#     bay_pumas %>% 
#       select(PUMACE10),
#     by = c("PUMA" = "PUMACE10")
#   ) %>% 
#   st_as_sf()
#   ) 
# ```
# 
# ```{r}
# bay_pums_couple <-
#   cleaned %>%
#   mutate(
#     WGTP = as.numeric(WGTP),
#     partner1 = ifelse(
#       (SPORDER == 1),
#       WGTP,
#       0
#     ),
#     partner2 = ifelse(
#       (SPORDER == 2),
#       WGTP,
#       0
#     ),
#     Same_sex_household = ifelse(
#       (HINCP < 100000) &
#         (CPLT == 2)|(CPLT == 4), 
#         
#       WGTP,
#       0
#     ),
#     Opposite_sex_household = ifelse(
#       (HINCP < 100000) &
#         (CPLT == 1)|(CPLT == 3),
#       WGTP,
#       0
#     ),
#     White = ifelse(
#       (SERIALNO == 1) &
#       (RAC1P == 1),
#       WGTP,
#       0
#     )
#     
# 
# 
# 
# 
# ```
# 
# 
# ```{r}
# acs_vars_2019_5yr <-
#   listCensusMetadata(
#     name = "2019/acs/acs5",
#     type = "variables"
#   )
# ```
# 
# 
# ```{r}
# veteran <-
#     getCensus(
#       name = "acs/acs5",
#       vintage = 2019,
#       region = "county:055",
#       regionin = "state:06",
#       vars = "group(S2101)" 
#     ) %>%
#       select(!c(GEO_ID,state,NAME,county) & !ends_with(c("EA","MA","M"))) %>%
#       pivot_longer(
#         ends_with("E"),
#         names_to = "name",
#         values_to = "estimate"
#       ) %>%
#       left_join(
#         acs_vars_2019_5yr %>% 
#           select(name, label)
#       )
# ```
# 
# ##How would we filter out just first two people to determine their races? Is sample size too small?
# 
# 
# ##Veterans, Jobs, Race< which war, income
# 
# 
# 
# 
# 
# 
# 






































